---
title: "Data 607 - Final Project"
output: html_document
---

### Project Proposal:

Source of motivation for the final project: Gentrification Analysis

NJ Housing Prices dataset: https://www.realtor.com/research/data/

Highest level of educational attainment:	Completing College

 https://data.ers.usda.gov/reports.aspx?ID=17829
 
Web Scraping list of fips code: https://en.wikipedia.org/wiki/List_of_United_States_FIPS_codes_by_county

API Call to census five years data:

https://api.census.gov/data/2019/acs/acs1?get=NAME,B01001_001E&for=county:*

I am sharing Forbes magazine's article. My source of motivation for this project is to investigate one of the paper’s key findings including:

"Students with a family income of $100,000 or more are more than twice as likely as students with family income under $50,000 to have combined SAT test scores of 1400 to 1600".

I will perform web scraping to fetch SAT data for several years and API call for the census five-year data. Both of my datasets will be for the state of New Jersey counties. I will perform the linear regression and try for other regression to find the correlation between the family income and students' SAT scores.

Also, perform the one of the machine learning to predict the next year's SAT outcome.


Importing Libraries

```{r}
library(tidycensus)
library(tidyverse)
library(dplyr)
library(plotly)
library(tidyr)
library(stringr)
library(rvest)

```

```{r}
train_df <- read_csv("https://raw.githubusercontent.com/uzmabb182/Data_607_Final_Project/main/2019loans.csv")

head(train_df)
```

```{r}
test_df <- read_csv("https://raw.githubusercontent.com/uzmabb182/Data_607_Final_Project/main/2020Q1loans.csv")

head(test_df)
```